%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Author : Jáchym Doležal, Tomas Polasek (template)
% Description : First exercise in the Introduction to Game Development course.
%   It deals with an analysis of a selected title from the point of its genre, 
%   style, and mechanics.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[a4paper,10pt,english]{article}

\usepackage[left=2.50cm,right=2.50cm,top=1.50cm,bottom=2.50cm]{geometry}
\usepackage[utf8]{inputenc}
\usepackage{hyperref}
\hypersetup{colorlinks=true, urlcolor=blue}

\newcommand{\ph}[1]{\textit{[#1]}}

\title{%
Analysis of Mechanics%
}
\author{%
Jáchym Doležal (xdolez0c)%
}
\date{}

\begin{document}

\maketitle
\thispagestyle{empty}

{%
\large

\begin{itemize}

\item[] \textbf{Title:} Escape From Tarkov

\item[] \textbf{Released:} 2017

\item[] \textbf{Author:} Battlestate Games, designed by Nikita Buyanov

\item[] \textbf{Primary Genre:} First-person shooter\



\item[] \textbf{Secondary Genre:} Extraction Shooter, Military Simulation, MMO, RPG

\item[] \textbf{Style:} Realistic Shooter, Immersive, Hardcore, 

\end{itemize}

}

\section*{\centering Analysis}

First-person shooters are among the most popular video game genres, with players increasingly seeking realism and intensity. Escape From Tarkov offers a unique FPS experience, known for its ability to immerse players. In EFT, players spawn on a map, loot, engage in combat with NPCs or other players, complete quests, and must either successfully extract or lose their gear upon death. EFT players can play different modes, such as PMC (Private Military Contractor), Scav (NPC or player-controlled scavenger), and offline. I'll focus on PMC gameplay for my analysis.

The FPS genre in EFT is reflected by realistic guns, realistic models of guns portraying real-world arms. Realistic simulation of bullets with a chance to ricochet. Together with large gun customization, EFT gunplay feels really realistic and immersive as some of their sounds have been recorded by the exact guns in the real world. The Map design is created to make extracting harder in certain scenarios. For example, on a map Reserve. PMCs can be extracted by using bunker complex D2. However, to be able to use this extraction point they have to pull a lever and go through a long corridor of bunkers, where there are tons of spots where other players can camp. A combination of all EFT gameplay mechanics it creates really unique experience each raid where players can feel a lot of adrenaline. Players get into really unpredictable situations, so they must be focused on the whole raid. 

On the other hand, EFT tries to balance realism and fun. Even though gunplay and some movement mechanics, like inertia walk and run speed are realistic, the jumping and health systems, are less realistic to create better PvP action and moments. The difficulty of the game is aimed at gunplay and low time to kill, so other mechanics are made less hardcore. For example, the healing system is simplified, so bleeding only decreases HP over time and there is no limit to how much blood they can lose.

MMO and RPG as sub-genres have some of their aspects in EFT. The PMC character stats and experience is the only thing players don't lose upon death. By playing PMC, players can complete quests which give them experience to their PMC level, and change the reputation of certain trades. With a higher PMC level and trader reputation. Players can unlock better gear they can buy from Traders. Players also can level their skills, like strength, stamina, metabolism and more. which for example decreases the chance of bleeding upon being hit. Besides character, players level up their skills with certain gun types, which decreases the recoil. These MMO and RPG elements give players a sense of progression which gives players the incentive to continue playing. Players feel a satisfying sense of growth and empowerment, even in the face of potential losses of wealth and equipment.

Tarkov's main style is immersion. Visually, The realistic graphics together with the gun and player animation make you really feel to be the character. Sounds mainly recorded by developers themselves help to create a sense of reality. For example, when eating a can of tuna the sounds are exactly as if you would eat it in real life. The other style of hardcore feel works well with immersion. As the raids feel hardcore, players must be focused so they are more immersed. Moreover, because the game allows people to loot, players with full backpacks of valuable loot and 30 minutes of their time in raid, feel the weight of their death. This makes players focused and it increases their will to survive. The combination of immersivity and hardcore gameplay helps EFT make the players feel like they are part of the world, but it also ignites an adrenaline rush that keeps them engaged and determined to survive.


\end{document}
